---
title: "Pharmacogenomic Analysis of Drug Combination Experiments to Identify
  Biomarkers of Response or Resistance"
author: 
  - name: Christopher Eeles
    email: christopher.eeles@uhnresearch.ca
    affiliation: 
        - &pm Princess Margaret Cancer Centre
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Pharmacogenomic Analysis of Drug Combination Experiments to Identify Biomarkers of Response or Resistance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Workshop Overview {-}

*This workshop is currently under development!*

## Instructor(s) name(s) and contact information {-}

* Christopher Eeles <christopher.eeles@uhnresearch.ca>
* Accepting volunteers! Who wants to help?

## Workshop Description {-}

This workshop will introduce users to the CoreGx and PharmacoGx R packages,
which are useful tools for pharmacogenomic modelling to discover biomarkers of
treatment response in cancer model systems. PharmacoGx specifically focuses on
drug sensitivity experiments in cancer cell lines, which will be the major
focus of this workshop. Additional infrastructure from our lab includes
ToxicoGx for toxicogenomics in healthy human cell-lines, RadioGx for
radiogenomics in cancer cell-lines and Xeva for pharmacogenomics in patient
derived xenograph (PDX) murine models.

Participants will learn the fundamentals of using CoreGx and PharmacoGx to
create a `PharmacoSet`&mdash;an integrative container for the storage, analysis and
visualization of pharmacogenomic experiments. Particular focus will be placed
on newly developed support for storing, analyzing and visualizing drug 
combination sensitivity experiments and correlating results therefrom for
with multi-omic molecular profiles to discover biomarkers of drug senstivity
or resistance.

### Pre-requisites {-}

* Basic knowledge of R syntax
* Knowledge of or interest in pharmacogenomics
* Familiarity with core Bioconductor classes such as the `SummarizedExperiment` and `MultiAssayExperiment`
* Curiosity or experience about `data.table` an assest!


Useful publications:

* Smirnov, P., Safikhani, Z., El-Hachem, N., Wang, D., She, A., Olsen, C., Freeman, M., Selby, H., Gendoo, D. M. A., Grossmann, P., Beck, A. H., Aerts, H. J. W. L., Lupien, M., Goldenberg, A., & Haibe-Kains, B. (2016). PharmacoGx: An R package for analysis of large pharmacogenomic datasets. Bioinformatics (Oxford, England), 32(8), 1244–1246. https://doi.org/10.1093/bioinformatics/btv723
* T.B.D. PharmacoGx Drug Combo Experiment Paper!


### Workshop Participation {-}

**No interactive portion yet!**

Participants are expected to have the following required packages installed
on their machines to be able to run commands allow with the instructors. ...

### _R_ / _Bioconductor_ packages used {-}

Bioconductor:

* https://www.bioconductor.org/packages/devel/bioc/html/CoreGx.html
* https://www.bioconductor.org/packages/devel/bioc/html/PharmacoGx.html

CRAN:

* https://cran.r-project.org/web/packages/data.table/index.html

## Time outline {-}

**Tentative, likely to change!**

For a 1-hr workshop:

| Activity                                                   | Time |
|------------------------------------------------------------|------|
| Introduction to CoreGx and PharmacoGx                      |  5m  |
| Generalizing Package Nomenclature                          |  5m  |
| Overview of Data Structures                                |  10m |
| New Features for Drug Combination Support                  |  15m |
| Building a Drug Combination PharmacoSet                    |  10m |
| Drug Combination Modelling (development in progress)       |  10m |
| Visualizing Drug Combination Experiments (brainstorming)   |  10m |

## Workshop goals and objectives {-}

T.B.D.

### Learning goals {-}

T.B.D.

### Learning objectives {-}

T.B.D.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
library(CoreGx)
library(PharmacoGx)
#library(PGxBioc2022Workshop)
```

# Introduction

This tutorial, titled Pharmacogenomic Analysis of Drug Combination Experiments 
to Identify Biomarkers of Response or Resistance, focuses on using the
`PharmacoGx` R package to correlate treatment response, measured as the
viability of cancer cell-lines after _in vivo_ drug treatment, with their
respective multi-omic profiles. `CoreGx` provides the core infrastructure
for storing, analyzing and visualizing generic treatment response experiments.
It provides functions and classes which can be inherited in downstream packages,
such as `ToxicoGx` and `RadioGx`. We hope that the `CoreSet` object is Generalizing
enough that it can be reused by other developers for their specific treatment
response use case.

# Generalizing Package Nomenclature

To facilitate modularization of the GxSuite of R packages, we are shifting
the nomenclature within a `CoreSet`&ndash;and therefore in inheriting packages
&ndash;to be more general.

To this end, we are making the following changes:

* Previous reference to cell (cell-line) will become sample, allowing the `CoreSet` to be used for other model systems
* Drug (radiation in `RadioGx`) will become treatment, allowing the `CoreSet` to be treatment type agnostic
* 

As a result of these changes, the names of some common accessors will be updated.
The old accessors will still remain, ensuring backwards compatibility for at
least two Bioconductor releases. A deprecation warning will be added to old 
accessors informing users of the corresponding new function, as per Bioconductor
best practices.


# Data Structures Overview

The GxSuite of packages make use of various Bioconductor classes for storing
molecular profile data. 

## CoreSet



## PharmacoSet



## TreatmentResponseExperiment


## MultiAssayExperiment


# New Features for Drug Combination Support



# Building a Drug Combiation PharmacoSet

## Raw Data


## Building a Treatment Response Experiment


## Building a MultiAssayExperiment


## Building the PharmacoSet


## Computing 


# Drug Combination Modelling

_Currently in development_

# Visiualizing Drug Combination Experiments

_Brainstorming session!_


# References




# SessionInfo

```{r session_info, include=FALSE}
sessionInfo()
```